---
title: "breakDown plots for the ranger model"
author: "Przemyslaw Biecek"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{breakDown plots for the ranger model}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Here we will use the HR churn data  (https://www.kaggle.com/ludobenistant/hr-analytics/data) to present the `breakDown` package for `ranger` models.

The data is in the `breakDown` package

```{r}
library(breakDown)
head(HR_data, 3)
```

Now let's create a `ranger` classification forest for churn, the `left` variable.

```{r}
library(ranger)
HR_data$left <- factor(HR_data$left)
model <- ranger(left ~ ., data = HR_data, importance = 'impurity', min.node.size = 10)
```

Variable importance for all trees in the forest.

```{r}
importance(model)
```

But how to understand which factors drive predictions for a single observation? 

With the `breakDown` package!

Explanations for the trees votings.

```{r, fig.width=7}
library(ggplot2)
explain_1 <- broken(model, HR_data[1159,])
explain_1
plot(explain_1) + ggtitle("breakDown plot for linear predictors of leaving the company")+ scale_y_continuous( limits = c(0,1), name = "fraction of trees", expand = c(0,0))

explain_1 <- broken(model, HR_data[10099,])
explain_1
plot(explain_1) + ggtitle("breakDown plot for linear predictors of leaving the company")+ scale_y_continuous( limits = c(0,1), name = "fraction of trees", expand = c(0,0))
```

This is not the right approach.
